Benchmarking on Rigetti architecture

### 2022-11-07 txt/maths.tex
add mathematical foundations

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{Benchmarking on Rigetti architecture}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \address[label1]{}
%% \address[label2]{}

\author{Paulina Lewandowska}
\author{\L ukasz Pawela}

\address{Institute of Theoretical and Applied Informatics, Polish Academy
	of Sciences, Ba{\l}tycka 5, 44-100 Gliwice, Poland}

\begin{abstract}
We introduce library ?, a comprehensive open source framework for benchmarking 
Amazon Braket Rigetti IonQ architectures ? 
The framework is built around  the scheme of discrimination of von Neumann 
measurements. The library provides algorithm for ... Our main aim is creating a 
common platform  for open research.  ?

\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
Benchmarking of Rigetti architecture \sep Discrimination of quantum 
measurements \sep Von Neumann measurements discrimination

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

\section*{Required Metadata}
\label{}

\section*{Current code version}
\label{}

Ancillary data table required for subversion of the codebase. Kindly replace 
examples in right column with the correct information about your current code, 
and leave the left column as it is.

\begin{table}[H]
\begin{tabular}{|l|p{6.5cm}|p{6.5cm}|}
\hline
\textbf{Nr.} & \textbf{Code metadata description} & \textbf{Please fill in this 
column} \\
\hline
C1 & Current code version & For example v42 \\
\hline
C2 & Permanent link to code/repository used for this code version & For 
example: $https://github.com/mozart/mozart2$ \\
\hline
C3 & Code Ocean compute capsule & For example: 
$https://codeocean.com/2017/07/30/neurospeech-colon-an-open-source-software-for-parkinson-apos-s-speech-analysis/code$\\
\hline
C4 & Legal Code License   & List one of the approved licenses \\
\hline
C5 & Code versioning system used & For example svn, git, mercurial, etc. put 
none if none \\
\hline
C6 & Software code languages, tools, and services used & For example C++, 
python, r, MPI, OpenCL, etc. \\
\hline
C7 & Compilation requirements, operating environments \& dependencies & \\
\hline
C8 & If available Link to developer documentation/manual & For example: 
$http://mozart.github.io/documentation/$ \\
\hline
C9 & Support email for questions & \\
\hline
\end{tabular}
\caption{Code metadata (mandatory)}
\label{} 
\end{table}

\linenumbers

%% main text

The permanent link to code/repository or the zip archive should include the 
following requirements:

README.txt and LICENSE.txt.

Source code in a src/ directory, not the root of the repository.

Tag corresponding with the version of the software that is reviewed.

Documentation in the repository in a docs/ directory, and/or READMEs, as 
appropriate.

\section{Motivation and significance}

Noisy intermediate-scale quantum
(NISQ)~\cite{preskill} devices are currently storming the market. We
have a selection of readily available quantum computers based on different
architectures. Historically, the first widely introduced quantum computing
architecture is D-Wave's quantum annealer~\cite{}.

https://quantumcomputingreport.com/scorecards/qubit-technology/

Additional approaches to quantum computing are currently in development. For 
instance we have Ionq's computer based on ion traps and Xanadu's optical 
lattices. As they are in prototype versions no access to general public is 
provided.

Next, we have computers implementing the gate model of quantum computation.
These are the best potentially developed machines, which can be thought
of as fully quantum computers, by which we understand that the qubits can be in
an entangled state.

Currently one of the main providers of such architectures is Rigetti and its 
Quantum Cloud Services platform. The company has
released three generations of QPUs so far: Rigetti 8Q Agave, Rigetti 19Q Acorn
and Rigetti 16Q Aspen-1. The most recent one was deployed in November 2018 and
is a downgrade in terms of number of qubits compared to the previous 19 qubit
one. However, the new chip is reportedly much more robust to noise and the
proposed architecture is much easier to scale up. The new unit is supposed to
serve as a building block for a 128 qubit
system~\footnote{https://medium.com/rigetti/the-rigetti-128-qubit-chip-and-what-it-means-for-quantum-df757d1b71ea}.
As was the case for IBM, Rigetti also provides a \texttt{Python} library, 
\texttt{pyquil}~\cite{}, which enables easy access to the machine.

\todo[inline]{co napisac o tym Amazonie, czy nic ?}

\todo[inline]{rysunki architektur}

%
%\begin{figure}[h]
%	\centering 
%	
%	\includegraphics[width=0.9\linewidth]{aspen.png} 
%	\caption{?}
%	\label{fig:aspen}
%\end{figure}

Now the natural question arises: how to construct a good metric of the
computational power of such devices? 
Lately, there have emerged propositions from the scientific community on how to
benchmark such devices.

 First let us mention the work by Michielsen
\emph{et.al.}~\cite{michielsen2017benchmarking}. In there the authors study the
IBM-QE device. Their first test is the creation of the entangled state
$\frac{1}{\sqrt{2}} (\ket{01} - \ket{10})$. Next they implement a
two-qubit+two-qubit adder. Another test is an identity operation, realized by an
even number of CNOT operations. Their final test is a quantum error correction
scheme (is it possible with measurement-controlled-operations?). The final
conclusion is that, except for simple circuits, the device fails to return
correct results.

Another, recent approach is utilizing quantum communication protocols in testing
of quantum architectures~\cite{zhukov2019quantum}. Here, the authors decide to
utilize superdense coding and the BB84 quantum cryptography protocol as
benchmarks for the IBM Q devices. They utilize the mutual information of the
transferred bits as a figure of merit for the quantum computer.

Other approaches focus on generative model training. One of these
works~\cite{hamilton2018generative} is aimed at small architectures, up to 5
qubits. Another recently presented is much more
robust~\cite{benedetti2018generative}. The main drawback of both of them is the
level of complication.

The goal of this work is to introduce a new benchmark for Rigetti 
architecture...

\label{}
\todo[inline]{
	Introduce the scientific background and the motivation for developing the 
	software. - motywacje

	Explain why the software is important, and describe the exact (scientific) 
	problem(s) it solves. - czemu ejst wazne, jakie problemy rozwiazuje

	Indicate in what way the software has contributed (or how it will 
	contribute in the future) to the process of scientific discovery; if 
	available, this is to be supported by citing a research paper using the 
	software. - do czego sie to przyczyni w przyszlosci

	Provide a description of the experimental setting (how does the user use 
	the software?). - w jaki sposób użytkownik korzysta z oprogramowania?

	Introduce related work in literature (cite or list algorithms used, other 
	software etc.). - inne oprogramowania i algorytmy }

\section{Software description}
\label{}
\todo[inline]{Describe the software in as much as is necessary to establish a 
vocabulary needed to explain its impact. }

We will first give a general overview of the structure of the code in 
Section~\ref{sec:sortware-architecture} and then provide additional details on 
the functionality of benchmark's schemes  in 
Section~\ref{sec:sortware-functionalities}.

\subsection{Software Architecture}\label{sec:sortware-architecture}
\label{}
\todo[inline]{
	Give a short overview of the overall software architecture; provide a 
	pictorial component overview or similar (if possible). If necessary provide 
	implementation details.}

\subsection{Software Functionalities}\label{sec:sortware-functionalities}
\todo[inline]{Present the major functionalities of the software.}

In the scope of this Paper, we aim at  exploring ideas revolving around quantum 
gate model-inspired computing devices, and assess their feasibility to 
benchmark a modern quantum architectures. The main idea that we have envisioned 
for this project is  introducing new concepts of benchmarks for NISQ devices 
benchmarking. This project aims to characterize the computing power and 
investigate possible practical applications of such devices having access by 
Amazon Braket.

\subsubsection{Model specification}
The scenario assumes that Alice and Bob have access to an unknown measurement
device, a black box. Alice has hidden one of the two measurements, either $\PP$
or $\mathcal{Q}$, in the black box. Bob only has the information that the black box
performs one of these measurements. They want to determine an upper bound on the
probability of correct discrimination between $\PP$ and $\mathcal{Q}$. Moreover, they also want to
construct an optimal discrimination strategy to achieve the highest probability. Theoretical results were characterized in~\cite{puchala2018strategies}.

In this work, we will consider the problem of discrimination between von Neumann 
measurements. One of them on the computational basis, the second one on basis 
$U$, where $U$ is a unitary matrix. 
The scheme of discrimination can be divided into the following steps, which are 
illustrated in Fig.~\ref{fig:theoretical_scheme}.

\begin{figure}[h!]
	\centering
	\includegraphics[scale=1.7]{pics/theoretical_scheme}
	\caption{Theoretical  scheme of discrimination  between von Neumann measurements $\PP_{U}$ and $\PP_\Id$. }
	\label{fig:theoretical_scheme}
\end{figure}

%\begin{figure}[h!]
%	\centering 
%	\includegraphics[width=0.9\linewidth]{gates1.pdf} 
%	\caption{A schematic representation of the setup for distinguishing
%		measurements by using entanglement. }
%	\label{fig:scheme}
%\end{figure}

\begin{enumerate}
	\item Alice and Bob share an entangled state $\ket{\psi_{AB}}$ on systems 
	$A$ and $B$.
	\item Alice performs one of two known measurements (either $\PP_{\Id}$ or 
	$\PP_{U}$) on part $A$ of the input state  $\ket{\psi_{AB}}$ and measures the system $A$.
	\item Alice performs the conditional binary measurement $\PP_{V_i}$ on the part 
	$B$ based on the output label $i$. 
	\item  Bob measures the system $B$ of the input state  $\ket{\psi_{AB}}$ and 
	makes a decision based on the received label $j$. If $j=0$, then he decides 
	that the the black box contains $\PP_U$. Otherwise, he decides that the 
	black box contains $\PP_{\Id}$. 
\end{enumerate}

@@

Let us take a closer look at the setup for distinguishing von Neumann 
measurements (see Fig.~\ref{fig:theoretical_scheme}). One of the components of 
this scheme is to perform a conditional binary measurement. 
Unfortunately, current NISQ devices do not allow for performing conditional 
binary measurements.
The only possible measurement performed on quantum architectures is the 
measurement in the computational basis. 
We  will present two possible methods providing a clever way to cross this 
obstacle. These will be two schemes of practical realization. 
The first method consists in using postselection while the second method 
consists in applying a controlled unitary. These methods are described below.

@@

\begin{enumerate}
\item Alice and Bob prepare an entangled state $\ket{\psi_{AB}}$ on systems 
$A$ and $B$.
\item Alice performs one of two quantum channels, either $\Phi_{\Id}$ or 
$\Phi_{U}$,  on part $A$ of the input state  $\ket{\psi_{AB}}$.
\item Alice measures part $A$ of the input state  $\ket{\psi_{AB}}$ and 
receives an output label $i$.
\item  
Based on Holevo--Helstrom theorem, Bob performs a conditional binary 
measurement	$\PP_{V_j^\dagger}$ on part $B$.
\item To calculate the probability of correct discrimination, they include just those cases for which $i = j$.
\item Bob measures part $B$ of the input state  $\ket{\psi_{AB}}$ and makes a
decision based on received label $j$. If $i=j=0$, then he decides that the 
black box contains $\PP_U$. Otherwise, he decides that the black box contains
$\PP_{\Id}$.
\end{enumerate}

@@

Due to the fact that the measurement result is nondeterministic, we are forced to consider the following circuits
in the postselection procedure
\begin{equation}
\begin{split}
(\Phi_{U}, \mathcal{P}_{V_k}, i, j) \\ 
(\Phi_{\Id}, \mathcal{P}_{V_k}, i, j)
\end{split}
\end{equation}
where $k=\{0,1\}, i=\{0,1\}$ and $ j=\{0,1\}$. 
The optimal probability of correct discrimination between $\PP_{U} $ and $\PP_\Id$ is as follows.
		\begin{equation}
		p = \frac{\#(\Phi_U, \PP_{V_i},i,0) + \#(\Phi_\Id, \PP_{V_i},i,1)}{\#(\Phi_U, \PP_{V_i},i,j) + \#(\Phi_\Id, \PP_{V_i},i,j)}, 
		\end{equation}
where $i=\{0,1\}$ and $ j=\{0,1\}$.

@@

\begin{enumerate}
\item Alice and Bob prepare an entangled state $\ket{\psi_{AB}}$ on 
systems $A$ and $B$.
\item Alice performs one of two quantum channels, either $\Phi_{\Id}$ or
$\Phi_{U}$,  on part $A$ of the input state  $\ket{\psi_{AB}}$.
\item Alice and Bob implement controlled unitary $V_0 \oplus V_1$ obtained from 
Holevo--Helstrom theorem.	
\item Alice and Bob prepare the measurement $\Delta$ in computational basis on 
their systems.
\item Bob makes a decision based on received label $j$. If $j=0$, then he
decides that the the black box contains $\PP_U$. Otherwise, he decides that the
black box contains $\PP_{\Id}$.
\end{enumerate}

@@

Here, we consider the following circuits
by using controlled unitary
\begin{equation}
\begin{split}
(\Phi_{U}, V_0 \oplus V_1, i, j) \\ 
(\Phi_{\Id},  V_0 \oplus V_1, i, j)
\end{split}
\end{equation}
where $i=\{0,1\}$ and $ j=\{0,1\}$. 
Then, the optimal probability of correct discrimination between $\PP_{U} $ and $\PP_\Id$ is as follows.
\begin{equation}
p = \frac{\#(\Phi_U,V_0 \oplus V_1,i,0) + \#(\Phi_\Id,V_0 \oplus V_1,i,1) }{\#(\Phi_U,V_0 \oplus V_1,i,j) + \#(\Phi_\Id,V_0 \oplus V_1,i,j)}
\end{equation}
where $i=\{0,1\}$ and $ j=\{0,1\}$. 
\end{scheme}

@@

Let $M_{d_1,d_2}$ be the set of all matrices of dimension $d_1 \times d_2$ over
the field $\mathbb{C}$. For  simplicity, square matrices will be denoted by
$M_d$.  The subset of $M_d$ consisting of Hermitian matrices of dimension $d$
will  be  denoted  by $\HH_d$,  while  the  set  of  positive semidefinite
matrices of dimension $d$ by $\HH_d^+$. The set of quantum states, that is
positive semidefinite operators having trace equal to one, will be denoted
$\Omega_d$. 
 The subset of $M_d$ consisting of unitary matrices will be denoted
by $\UU_d$, while its subgroup of diagonal unitary operators will be denoted by
$\DD \UU_d$. 
	An operator $A = \left( a_{i,j}\right)_{i,j} \in M_d$ is said to be a stochastic matrix if $a_{i,j} \ge 0$   and $\sum_{i} a_{i,j} = 1$.
Whereas an operator $A = \left( a_{i,j}\right)_{i,j} \in M_d$ is said to be a double stochastic matrix if $A$ is a stochastic matrix and $\sum_{j} a_{i,j} = 1$.

 We will also need a linear mapping transforming $M_{d_1}$ into
$M_{d_2}$, which will be denoted 
\begin{equation}
\Phi: M_{d_1 } \rightarrow M_{d_2}.
\end{equation} 
There
exists a bijection between introduced linear mappings $\Phi$ and set of matrices $M_{d_1d_2}$,  known as the Choi-
Jamio{\l}kowski isomorphism. 
For a given linear mapping $\Phi_M$ corresponding  Choi operator $\mathcal{J}(\Phi)$ is explicitly written as 
\begin{equation}
\mathcal{J}(\Phi) \coloneqq \sum_{i,j=1}^{d- 1} \Phi(\ketbra{i}{j}) \otimes \ketbra{i}{j}. \end{equation}
Finally, we introduce a special subset of all mappings $\Phi$, called quantum channels, which are completely positive
and trace preserving (CPTP).
In this work we will consider a special class of quantum channels, called unitary channels. Let $U \in \UU_d$ be a unitary operator. A unitary channel
$\Phi_{U}$ is defined as $\Phi_U(\cdot) = U \cdot U^\dagger$ for  $U \in 
\UU_d$. A general quantum
measurement, that is a positive operator valued measure (POVM) $\PP$ is a
collection of positive semidefinite operators $\{E_1, \ldots, E_m \}$ called
\emph{effects}, which sum up to identity, \ie $ \, \, \sum_{i=1}^m E_i = \1$. If
all the effects are rank-one projection operators, then such a measurement is
called a von Neumann measurement. Every von Neumann measurement can be
parameterized by a unitary matrix and hence we will use the notation $\PP_{U}$
for a von Neumann measurement with effects $\{\proj{u_0}, \ldots, \proj{u_{d-1}}\}$,
where $\ket{u_i}$ is the $i+1$-th column of the unitary matrix $U$. The action of
quantum measurement $\PP_{U}$ on some state $\rho \in \Omega_d$ can be
seen as  a measure-and-prepare quantum channel as follows 
$
\PP_{U} : \rho \rightarrow \sum_{i=0}^{d-1} \bra{u_i} \rho \ket{u_i} \proj{i}.
$

@@

Let us  consider a  linear mapping transforming  square matrices into square 
matrices that is $\Phi: M_{d_1} \to M_{d_2}$. 
We define the diamond norm of a map $\Phi$ as 
\begin{equation}
\|\Phi\|_\diamond = \max_{\|X\|_1=1} \| \left(\Phi \otimes \1\right) (X) \|_1.
\end{equation}

The celebrated result by Helstrom~\cite{helstrom1976quantum} gives the optimal  probability of correct discrimination between two quantum operations, $\TT$  and $\SS$, 
in terms of their distance with the use of the diamond norm
\begin{equation}
p_{\text{succ}}(\TT, \SS) =  \frac12 + \frac14 \| \TT - \SS \|_\diamond.
\end{equation}
In the case of discrimination between two von Neumann measurements $\PP_U$ and 
$\PP_\Id$ we know that (from Theorem 1 in \cite{puchala2018strategies})
\begin{equation}
\|\PP_U - \PP_\Id\|_\diamond = \min_{E \in \diaguni_d} \|\Phi_{UE} - 
\Phi_\Id\|_\diamond,
\end{equation}
where $\Phi_U$ is unitary channel and $\diaguni_d$ be the set of diagonal 
unitary matrices of dimension $d$.

\paragraph{Discrimination of unitary channels}

Before we proceed to presenting our main results, we need to briefly discuss the
problem of discrimination of unitary channels.  For this purpose, we introduce the notion of numerical range of
a given matrix $A \in M_d$. The set \begin{equation}
W(A) =\{\bra{x}A\ket{x}: \ket{x} \in 
\mathbb{C}^d, \;
\;\braket{x}{x}=1\}.
\end{equation}   
is the numerical range of $A$.
The detailed properties of the numerical range and its generalizations we can see on the website~\cite{nr}.

 Let us now recall the well-known result for the
distinguishability of unitary channels.
	Let $U \in \UU_d$ and $\Phi_U$ be a unitary 
	channel. 
	Then, the following equation holds
	\begin{equation}
	\| \Phi_U  - \Phi_{\1} \|_\diamond = 2 \sqrt{1-\nu^2},
	\end{equation}
	where $\nu = \min_{x \in W(U^\dagger)} |x|  $. And hence the optimal probability to discriminate two unitary channel is given by 
	\begin{equation}
	p_{\text{succ}}(\Phi_U, \Phi_\Id) = \frac{1}{2} + \frac{1}{2} \sqrt{1-\nu^2}.
	\end{equation}

%In the general case, the diamond norm of a Hermiticity-preserving maps $\Phi$ 
%can be computed using the Semidefinite Program~\cite{watrous2018theory} and 
%state 
%primal and dual problem in the following form:

%
%\begin{minipage}{0.495\linewidth}
%	\begin{equation*}
%		\begin{split}
%			\text{\textbf{Primal problem:}} \\
%			\text{maximize:}\quad & \tr(X J(\Phi)) \\[2mm]
%			\text{subject to:}\quad &  \left[ \begin{array}{cc}\Id_{d_2} 
%\otimes \rho & X \\ X^* & \Id_{d_2} \otimes \rho  \end{array} \right] \ge 0 \\
%			& \rho \in \HH_d^+\\ 
%			& X \in M_{d1,d2}.
%		\end{split}
%	\end{equation*}
%\end{minipage}
%\begin{minipage}{0.495\linewidth}
%	\begin{equation*}
%		\begin{split}
%			\text{\textbf{Dual problem:}} \\
%			\text{minimize:}\quad & ||\tr_1(Y)||_\infty \\[2mm]
%			\text{subject to:}\quad & \left[ \begin{array}{cc}Y & -J(\Phi) \\ 
%-J(\Phi) &  Y \end{array} \right] \ge 0 \\
%			& Y \in \HH_{d_1d_2}^+.
%		\end{split}
%	\end{equation*}
%	\vspace{0.5cm}
%\end{minipage}

@@

Let us focus on single-qubit von Neumann measurements $\PP_\1$ and $\PP_U$.
Assume that the unitary matrix $U$ is of the form 
\begin{equation}
U = H 
\left(\begin{array}{cc}1&0\\0&e^{i \phi}\end{array}\right)  H^\dagger
\end{equation}
%\begin{equation}
%U = H \diag (1, \ee^{\ii \phi}) H^\dagger,
%\end{equation}
where $H$ is the Hadamard matrix of dimension two and $\phi \in [0, 2 \pi)$.
In this section we present theoretical probability of correct 
discrimination between these measurements. Moreover, we create the optimal 
theoretical strategy of their discrimination. 
The otimal probability will be calculated in 
subsection~\ref{sec:example_theoretical_probability}. Later, we will 
discuss the realization of this scheme in 
subsection~\ref{sec:example_realization}. For such realization, we 
will need the proper input state, (see sec.~\ref{sec:example_discriminator}) 
and final measurement (see sec.~\ref{sec_example_final_measurement}).

\subsection{Theoretical probability}\label{sec:example_theoretical_probability}

In this section we calculate the theoretical probability of correct discrimination between two von Neumann measurements $\PP_\1$ and $\PP_U$ for $U = H \diag(1, e^{i \phi}) H^\dagger$, where  $\phi \in [0, 2\pi)$.  To do that, we will present an auxiliary lemma.   
\begin{lemma}\label{lemma:min-e-optimal}
Let $U = H \diag(1, e^{i \phi}) H^\dagger$, $\phi \in [0, 2\pi)$ and	let 
$\Phi_U$ and $\Phi_\Id$ be two unitary channels. Then, the following equation holds 
\begin{equation}
\min_{E \in \diaguni_2} \|\Phi_{UE} - 
\Phi_\Id\|_\diamond = \|\Phi_{U} - 
\Phi_\Id\|_\diamond,
\end{equation}
\end{lemma}

\begin{proof} Recall that the distance between two unitary channels is given by
	$
	\| \Phi_U  - \Phi_{\1} \|_\diamond = 2 \sqrt{1-\nu^2},
	$
	where $\nu = \min_{x \in W(U^\dagger)} |x|  $ for any $U \in \mathcal{U}_d$. 
	For $U = H 
	\left(\begin{array}{cc}1&0\\0&e^{i \phi}\end{array}\right)  H^\dagger$ the readers briefly observe that  $\nu^2 = 1 - \frac{|1 - e^{-i \phi} |^2 }{4} = 1 - \frac{|1 - e^{i \phi} |^2 }{4}$. So, 
	\begin{equation}
	\|  \Phi_U  - \Phi_{\1} \|_\diamond = | 1 - e^{i \phi} |. 
	\end{equation} 
		It implies that it is enough to prove  \begin{equation}
		\min_{E \in \diaguni_2} \|\Phi_{UE} - 
		\Phi_\Id\|_\diamond  = | 1 - e^{i \phi} |.
		\end{equation}
%		It implies that we can prove  equivalently  the following condition
%		\begin{equation}
%		\max_{E \in \diaguni_2 } \nu_{UE} = \nu_U
%		\end{equation}
This condition is equivalent to 
	\begin{equation}
	\max_{E \in \diaguni_2 } \nu_{E} = \frac{|1 + e^{i \phi} | }{2},
	\end{equation}
	where $\nu_E = \min_{x \in W(U^\dagger E)} |x|. $

>>>>

\subsubsection{Model specification}
The scenario of measurements discrimination assumes that we have access to an unknown measurement
device. One of the two measurements, either $\PP$
or $\mathcal{Q}$,  was used during an experiment. The only information we have that one of these measurements was 
performed. We want to determine an upper bound on the
probability of correct discrimination between $\PP$ and $\mathcal{Q}$. Moreover, we also want to
construct an optimal discrimination strategy to achieve the highest possible probability.

Discriminating between von Neumann measurements is valuable from the applications viewpoint. That is because, unlike other types of measurements, they can be implemented on current hardware.
So
here, we will consider the discrimination task between two von Neumann measurements, one of them on the computational basis, whereas the second one on basis 
$U$, where $U$ is a unitary matrix. 
Theoretical scenario was characterized in~\cite{puchala2018strategies}.
Such a scheme can be best described into the following steps 
and illustrated in
Fig.~\ref{fig:theoretical_scheme}.

\begin{figure}[h!]
	\centering
$  $	\includegraphics[scale=1.7]{pics/theoretical_scheme}
	\caption{Theoretical  scheme of discrimination  between von Neumann measurements $\PP_{U}$ and $\PP_\Id$. }
	\label{fig:theoretical_scheme}
\end{figure}

\begin{enumerate}
	\item We initialize the discriminator $\ket{\psi_{0}}$ on bipartite  system.
	\item We perform one of two known measurements (either $\PP_{\Id}$ or 
	$\PP_{U}$) on first part of the input state  $\ket{\psi_{0}}$, that means $\left(\PP \otimes \Id \right) (\proj{\psi_0})$.  
	\item We measure the  first system and obtain a label $i$. 
	\item We perform the conditional binary  (optimal) measurement $\PP_{V_i}$ on the second part based on the output label $i$. 
	\item  We measure the second  system of the input state  $\ket{\psi_{0}}$ and 
	make a decision based on the received label $j$. If $j=0$, then we decide 
	that the $\PP_U$ occurs. Otherwise, we decide that $\PP_{\Id}$ occurs. 
\end{enumerate}

@@

Let us take a closer look at the setup for distinguishing von Neumann 
measurements (see Fig.~\ref{fig:theoretical_scheme}). One of the components of 
this scheme is to perform a conditional  measurement. 
However, current NISQ devices do not allow for performing conditional 
 measurements.

We  will present two possible methods providing a clever way to cross this 
obstacle. These will be two schemes of practical realization. 
The first method consists in using postselection while the second method 
consists in applying a controlled unitary. These methods are described below.

@@

\begin{enumerate}
\item We prepare  the discriminator $\ket{\psi_{0}}$ on bipartite system.
\item We perform one of two quantum channels, either $\Phi_{\Id}$ or 
$\Phi_{U^\dagger}$,  on  the first part of the input state  $\ket{\psi_{0}}$, that means $\left( \Phi \otimes \Id\right)(\proj{\psi_0})$.
\item We measure the first part of the input state  $\ket{\psi_{0}}$ and 
receive an output label $i$.
\item  
Based on Holevo--Helstrom theorem, we perform a conditional binary 
measurement	$\PP_{V_j}$ and hence, we implement the quantum channel $\Phi_{V_j^\dagger}$ on the second part. 
\item To calculate the probability of correct discrimination, we include just those cases for which $i = j$.
\item We measure the second part of the input state  $\ket{\psi_{0}}$ and make a
decision based on received label $j$. If $i=j=0$, then we decide that  
$\PP_U$ occurs.  If $i=j=1$, we decide that 
$\PP_{\Id}$ occurs.
\end{enumerate}

@@

Due to the fact that the measurement result is nondeterministic, we are forced to consider the following circuits
in the postselection procedure
\begin{equation}
\begin{split}
(\Phi_{U^\dagger}, \Phi_{V_k^\dagger}, i, j) \\ 
(\Phi_{\Id}, \Phi_{V_k^\dagger}, i, j)
\end{split}
\end{equation}
where $k=\{0,1\}, i=\{0,1\}$ and $ j=\{0,1\}$. 
The optimal probability $p_{\text{succ}}(\PP_{U}, \PP_{\Id}) $ of correct discrimination between $\PP_{U} $ and $\PP_\Id$ is as follows
		\begin{equation}
		p_{\text{succ}}(\PP_{U}, \PP_{\Id})= \frac{\#(\Phi_{U^\dagger}, \Phi_{V_i^\dagger},i,0) + \#(\Phi_\Id, \Phi_{V_i^\dagger},i,1)}{\#(\Phi_{U^\dagger}, \Phi_{V_i^\dagger},i,j) + \#(\Phi_\Id, \Phi_{V_i^\dagger},i,j)}, 
		\end{equation}
where $i=\{0,1\}$ and $ j=\{0,1\}$.

@@

\begin{enumerate}
\item We prepare the discriminator $\ket{\psi_{0}}$ on bipartite 
system. 
\item We perform one of two quantum channels, either $\Phi_{\Id}$ or
$\Phi_{U^\dagger}$,  on  the first part of the input state  $\ket{\psi_{0}}$, that means $\left(\Phi \otimes \Id\right)(\proj{\psi_0})$.
\item Based on Holevo--Helstrom theorem, we perform a conditional binary measurement $\PP_{V_j}$, and hence, we implement the direct sum of quantum channels $\Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger}$. Observe that   $\Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger} = \proj{0} \otimes \Phi_{V_0^\dagger}  + \proj{1} \otimes \Phi_{V_1^\dagger}.  $
\item We measure all systems in computational basis $\Delta$.
\item We make a decision based on received label $j$ on the second system. If $j=0$, then we
decide that $\PP_U$ occurs. Otherwise, we decide that $\PP_{\Id}$ occurs.
\end{enumerate}

@@

Here, we consider the following circuits
by using controlled unitary
\begin{equation}
\begin{split}
(\Phi_{U^\dagger}, \Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger}, i, j) \\ 
(\Phi_{\Id},  \Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger}, i, j)
\end{split}
\end{equation}
where $i=\{0,1\}$ and $ j=\{0,1\}$. 
Then, the optimal probability of correct discrimination between $\PP_{U} $ and $\PP_\Id$ is as follows.
\begin{equation}
p = \frac{\#(\Phi_{U^\dagger},\Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger},i,0) + \#(\Phi_\Id,\Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger},i,1) }{\#(\Phi_{U^\dagger},\Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger},i,j) + \#(\Phi_\Id,\Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger},i,j)}
\end{equation}
where $i=\{0,1\}$ and $ j=\{0,1\}$. 
\end{scheme}

@@

In this and the following sections we will denote complex Euclidean spaces $\mathbb{C}^d$ with $\XX, \mathcal{Y}$ etc. When needed the dimension of a space $\XX$ will be denoted $\dim(\XX)$. The set of matrices transforming vectors from $\XX$ to $\mathcal{Y}$ will be denoted $\mathrm{L}(\XX, \mathcal{Y})$.
 We will also need a linear mapping transforming $\mathrm{L}(\XX)$ into
$\mathrm{L}(\mathcal{Y})$, which will be denoted 
\begin{equation}
\Phi: \mathrm{L}(\XX) \rightarrow \mathrm{L}(\mathcal{Y}).
\end{equation} 
There
exists a bijection between introduced linear mappings $\Phi$ and set of matrices $\mathrm{L}(\mathcal{Y} \otimes \mathcal{X})$,  known as the Choi-
Jamio{\l}kowski isomorphism. 
For a given linear mapping $\Phi$ the corresponding  Choi operator $J(\Phi)$ is explicitly written as 
\begin{equation}
J(\Phi) \coloneqq \sum_{i,j=0}^{d- 1} \Phi(\ketbra{i}{j}) \otimes \ketbra{i}{j}. \end{equation}
Finally, we introduce a special subset of all mappings $\Phi$, called quantum channels, which are completely positive
and trace preserving (CPTP).
We will consider a special class of quantum channels, called unitary channels. A unitary channel
$\Phi_{U}$ is defined as $\Phi_U(\cdot) = U \cdot U^\dagger$ for  and unitary $U$.

A general quantum
measurement, that is a positive operator valued measure (POVM) $\PP$ is a
collection of positive semidefinite operators $\{E_1, \ldots, E_m \}$ called
\emph{effects}, which sum up to identity, \ie $ \, \, \sum_{i=1}^m E_i = \1$. If
all the effects are rank-one projection operators, then such a measurement is
called a von Neumann measurement. Every von Neumann measurement can be
parameterized by a unitary matrix $U$ which the effects $\{\proj{u_0}, \ldots, \proj{u_{d-1}}\}$,
ale creating by taking $\ket{u_i}$ as  $i+1$-th column of the unitary matrix $U$. 
Hence, we will use the notation $\PP_{U}$. 
%
%
% The action of
%quantum measurement $\PP_{U}$ on some state $\rho \in \Omega_d$ can be
%seen as  a measure-and-prepare quantum channel as follows 
%$
%\PP_{U} : \rho \rightarrow \sum_{i=0}^{d-1} \bra{u_i} \rho \ket{u_i} \proj{i}.
%$

@@

We define the diamond norm of a map $\Phi$ as 
\begin{equation}
\|\Phi\|_\diamond = \max_{\| \ket{\psi}\|_1=1} \| \left(\Phi \otimes \1\right) (\proj{\psi}) \|_1.
\end{equation}
The quantum state $\ket{\psi}$ which maximizes the diamond norm we will call as discriminator.

The celebrated result by Helstrom~\cite{helstrom1976quantum} gives the optimal  probability of correct discrimination between two quantum measurements, $\PP$  and $\mathcal{Q}$, 
in terms of their distance with the use of the diamond norm
\begin{equation}
p_{\text{succ}}(\PP, \mathcal{Q}) =  \frac12 + \frac14 \| \PP - \mathcal{Q} \|_\diamond.
\end{equation}

@@

Let us focus on single-qubit von Neumann measurements $\PP_\1$ and $\PP_U$.
Assume that the unitary matrix $U$ is of the form 
\begin{equation}
U = H 
\left(\begin{array}{cc}1&0\\0&e^{i \phi}\end{array}\right)  H^\dagger
\end{equation}
%\begin{equation}
%U = H \diag (1, \ee^{\ii \phi}) H^\dagger,
%\end{equation}
where $H$ is the Hadamard matrix of dimension two and $\phi \in [0, 2 \pi)$.

\subsubsection{Probability}\label{sec:example_probability}
\begin{theorem}
The optimal probability of correct discrimination between von Neumann
measurements $\PP_U$ and $\PP_{\Id}$ is given by
\begin{equation}
p_{\text{succ}}(\PP_{U}, \PP_{\Id}) = \frac{1}{2} + \frac{|1 - e^{i \phi}  |}{4} . 
\end{equation}
\end{theorem}
